The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-03-19 paper/iphas-dr2.tex
Incorporating feedback from V.

\section{Introduction}
The INT/WFC Photometric H$\alpha$ Survey
of the Northern Galactic Plane \citep[IPHAS;][]{Drew2005}
is providing new insights into the contents and structure 
of our own backyard, the Milky Way.
The original motivation for undertaking 
this large-scale programme of observation
-- spanning almost a decade 
and using more than 300 nights 
at the Isaac Newton Telescope (INT) in La Palma -- 
was to provide the digital update 
to the photographic northern H$\alpha$ surveys 
of the mid-20th century. 
By increasing the sensitivity 
with respect to these previous surveys 
by a factor $\sim$1000 (7 magnitudes), 
it was envisaged that IPHAS would allow 
the limited bright samples of Galactic emission line objects 
available at the outset \citep[e.g.][]{Kohoutek1999}, 
to be extended into larger, deeper, more statistically-robust samples.
This in turn could better inform our understanding 
of the early and late stages of stellar evolution.
Since the publication of the IPHAS Initial Data Release \citep[IDR;][]{Gonzalez-Solares2008},
these aims have begun to be realised through a
range of published studies including: 
a preliminary catalogue of candidate emission line objects \citep{Witham2008};
discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
new cataclysmic variables \citep{Witham2007}; 
new groups of young stellar objects
\citep{Vink2008,Barentsen2011a,Wright2012};
%new asymptotic giant branch stars \citep{Wright2008};
new classical Be stars \citep{Raddi2013};
along with discoveries of new and remarkable planetary nebulae 
\citep{Mampaso2006, Corradi2011, Viironen2011}
and new supernova remnants \citep{Sabin2013}.

Over the years it has become apparent that the legacy of IPHAS 
will extend beyond the traditional \ha\ applications 
of identifying emission-line stars and nebulae. 
Through the provision of $r$, $i$ broadband photometry 
alongside H$\alpha$ data,
IPHAS has created the opportunity 
to study Galactic Plane populations 
in a new way.
For example, the survey’s unique $(r-\ha)$ colour, 
when combined with $(r-i)$,
has been shown to provide simultaneous constraints 
on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
This has opened the door 
to a wide range of Galactic science applications, 
including the mapping of extinction across the Plane in three dimensions
and the probabilistic inference of stellar properties
\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
In effect, the availability of narrowband H$\alpha$ 
alongside $r, i$ magnitudes
provides coarse spectral information for huge samples of stars 
which are otherwise too faint or numerous 
to be targeted by spectroscopic surveys.
For such science applications to succeed however, 
it is vital that the imaging data is transformed 
into a homogeneously calibrated photometric catalogue, 
in which quality problems 
and duplicate detections are flagged.

>>>>

\section{Introduction}
The INT/WFC Photometric H$\alpha$ Survey
of the Northern Galactic Plane \citep[IPHAS;][]{Drew2005}
is providing new insights into the contents and structure 
of our own backyard, the Milky Way.
The original motivation for undertaking 
this large-scale programme of observation
-- spanning almost a decade 
and using more than 300 nights 
at the Isaac Newton Telescope (INT) in La Palma -- 
was to provide the digital update 
to the photographic northern H$\alpha$ surveys 
of the mid-20th century. 
By increasing the sensitivity 
with respect to these previous surveys 
by a factor $\sim$1000 (7 magnitudes), 
it was envisaged that IPHAS would allow 
the limited bright samples of Galactic emission line objects 
available at the outset \citep[e.g.][]{Kohoutek1999}, 
to be extended into larger, deeper, more statistically-robust samples.
This in turn could better inform our understanding 
of the early and late stages of stellar evolution.
Since the publication of the IPHAS Initial Data Release \citep[IDR;][]{Gonzalez-Solares2008},
these aims have begun to be realised through a
range of published studies including: 
a preliminary catalogue of candidate emission-line objects \citep{Witham2008};
discoveries of new symbiotic stars \citep{Corradi2008, Corradi2010, Corradi2011}; 
new cataclysmic variables \citep{Witham2007}; 
new groups of young stellar objects
\citep{Vink2008,Barentsen2011a,Wright2012};
%new asymptotic giant branch stars \citep{Wright2008};
new classical Be stars \citep{Raddi2013};
along with discoveries of new and remarkable planetary nebulae 
\citep{Mampaso2006, Viironen2009a, Viironen2009b, Corradi2011, Viironen2011}
and new supernova remnants \citep{Sabin2013}.

Over the years it has become apparent that the legacy of IPHAS 
will also reach beyond these traditional \ha\ applications 
of identifying emission-line stars and nebulae. 
Through the provision of $r$, $i$ broadband photometry 
alongside H$\alpha$ data,
IPHAS has created the opportunity 
to study Galactic Plane populations 
in a new way.
For example, the survey’s unique $(r-\ha,\ r-i)$ colour-colour
diagram has been shown to provide simultaneous constraints 
on intrinsic stellar colour and interstellar extinction \citep{Drew2008}. 
This has opened the door 
to a wide range of Galactic science applications, 
including the mapping of extinction across the Plane in three dimensions
and the probabilistic inference of stellar properties
\citep{Sale2009, Sale2010, Giammanco2011, Sale2012, Barentsen2013}. 
In effect, the availability of narrowband H$\alpha$ 
alongside $r, i$ magnitudes
provides coarse spectral information for huge samples of stars 
which are otherwise too faint or numerous 
to be targeted by spectroscopic surveys.
For such science applications to succeed however, 
it is vital that the imaging data is transformed 
into a homogeneously calibrated photometric catalogue, 
in which quality problems 
and duplicate detections are flagged.
